---
title: Visualizing a Web of Citations Using the Google Scholar API and D3.js
author: Sebastian Schweer
date: '2018-01-09'
slug: visualizing-a-web-of-citations
categories:
  - R
  - D3.js
tags:
  - Visualization
  - APIs
---

> "I am very glad you asked me that, Mrs. Rawlinson. The term 'holistic' refers to my conviction that what we are concerned here is the fundamental interconnectedness of things."
Douglas Adams, "Dirk Gently's Holistic Agency"

Networks are *everywhere*. The internet, road infrastructure and even [the cast list of a theater show](d3 js link): depending on the interpretation, all kinds of entities can be viewed as networks. 

One particularly interesting subset of networks is made up by scientific research papers. For instance, the Erdös number is calculated using a network of authors by co-authorship: If you have co-written a paper with Paul Erdös: your Erdös number is 1. Else, your Erdös number is one higher than the smallest Erdös number of any of your co-authors. If you have never written a mathematical paper, your Erdös number is, unfortunately, infinity. 

A similar measurement was introduced for people starring in movies by calculating the Bacon-number, the "distance" to Kevin Bacon. Finally, the Bacon-Erdös-Number is reserved for the (rather elite) subgroup of people on this earth who have both
* Published a scientific paper and
* Starred in a movie.
Apparently, the lowest Bacon-Erdös number is being held by YYY. But I digress.

Another possibility of creating a network from scientific articles is looking at the citations: each paper that cites another paper creates a link between these articles. In this blog post I demonstrate how to

1. Obtain the connected graph between a set of articles by using the `rCrossRef` package
2. Visualize the results using D3.js 

# Get Data from CrossRef

In my days at the university, I managed to publish a handful of articles and I will use these as demonstrators in this post. First, let's install and load the library and use it to find all articles under my name:
```{r}
library("rcrossref")
articles_df <- cr_works(flq = c(query.author="Sebastian Schweer"))$data
articles_df[ , c("title", "DOI", "container.title", "created")]
```
Right, so apparently, I'm not the only "Sebastian Schweer" in the scientific world. In order to find my own works, we have to filter the entries a bit. I did my PhD research during the years 2011 until 2015, and some articles just take their time to be published. So let's just look at articles from the year 2014 until today. We find
```{r}
articles_df <- articles_df[ substr(articles_df$created, 1, 4) %in% c("2014","2015","2016","2017") , c("title", "DOI", "container.title", "created")]
```